
\documentclass[10pt]{article}
%\usepackage{tex4ht}
\usepackage[OT1,T2A]{fontenc}
\usepackage[koi8-u]{inputenc}
\usepackage{graphicx}
\usepackage{verbatim}

%\Contribute{TITLE}{\def\LaTeX{LaTeX}}
\title{ DirectoryContainer : Programmer Guide
        \newline
        \small{DocumentId:GradSoft-PR-e-11.12.2000-v1.0.0}
}
% $Id: ProgrammingGuide_eng.tex,v 1.6 2002-10-16 10:30:02 rssh Exp $

\begin{document}

\maketitle{}

\tableofcontents

\section{Introduction}

 This package DirectoryContainer is a cross-platform C++ component for reading file directory.
 % check this turn of speach is correct
 Using DirectoryContainer, you receives a simple API for actions such as searching file,
 the same for UNIX and for Windows 32.
 The package is supplied as source code.
 The package is created and suppported by GradSopft company,
 the home page of GradSoft is http://www.gradsoft.kiev.ua/.

  This document is non-formal description of package, full specification are in
  API reference. (www.gradsoft.kiev.ua/common/ToolBox/DirectoryContainer/API/).

\section{Principles of using}


\subsection{General information}

Two next approach to the reading of the directory is realized:
\begin{enumerate}
\item You can use a {\bf DirectoryEntry} object for handling directory like an input stream;
\item You can use a {\bf DirectoryContainer} object for handling directory like a container
      for elements of type DirectoryEntry.
\end{enumerate}

\subsection{Class DirectoryEntry}

\subsubsection{General information} 

DirectoryEntry is a {\em being-positioned} object for handling the directory like an input stream.
These words means following:
\begin{enumerate}
\item \begin{enumerate}
      \item directory is considered to be a sequence of some (specific) elemets,
            access to which is possible by unique pointer to the "current" element;

      \item DirectoryEntry is designed to be a mean of handling of a single (the current) element
            which is replaceable, however, by own methods of DirectoryEntry.
      \end{enumerate}

\item In order to directory was readed,
      that must be opend before and closed afterwards.
\end{enumerate}
Using DirectoryEntry, the directory claimed will be opend along with object creation, 
so that to obtain information about certain file you must fulfill the next actions:
\begin{enumerate}
\item To create DirectoryEntry and to {\em position} it at the file of interest;
\item To invoke methods which returns information about current element of sequence.
\end{enumerate}

\noindent Example:

\begin{verbatim}
try{

  DirectoryEntry smth(".");  // create object and position it
                             // at the first element of sequence
  do {

   cout<<smth.name()<<endl;  // display the name of element

  } while(smth.next());      // position object at the following 
                             // element of sequence (or break the loop)

} catch(DirectoryException& ex) {   // see below

  cerr << ex.message << endl;
}
\end{verbatim}
This code display the list of files disposed in the current directory.

\subsubsection{API}

{\bf General information:} \\

\noindent DirectoryEntry complies with models Assignable and EqualityComparable, i.e. it has:
\begin{enumerate}
\item Copy constructor and operator= ;
\item operator== and operator!=.
\end{enumerate}
Nevertheless, it is'nt DеfaultConstructable, default constructor of it is private.

\paragraph{Description of methods:}

\begin{enumerate}
\item Means for complying with Assignable and EqualityComparable 

      \begin{itemize}
      \item {\em DirectoryEntry(const DirectoryEntry\& )};
      \item {\em DirectoryEntry\& operator=(const DirectoryEntry\& )};
      \item {\em bool operator==(const DirectoryEntry\& ) const};
      \item {\em bool operator!=(const DirectoryEntry\& ) const};
      \end{itemize}

\item Means for creation and closing of stream:

      \begin{itemize}
      \item {\em DirectoryEntry(const char*)} - open directory and position object
            at the first element

            {\bf Parameter:} name of directory

      \item {\em close()} - close directory 
      \item {\em \verb|~|DirectoryEntry()} - invoke close()

      \end{itemize}

\item Means for positioning:

      \begin{itemize}
      \item {\em bool next()} - position object at the next element of sequence 

            {\bf Return value:}
            \begin{list}{-}{}
            \item {\bf true}, if positioning succeeds;
            \item {\bf false}, if object arrives at the end of file 
            \end{list}

      \item {\em bool find(const *char)} - position object at the element with 
            given name

            {\bf Parameter:}

            the name of the file must be found

            {\bf Return value:}
            \begin{list}{-}{}
            \item {\bf true}, if positioning succeeds;
            \item {\bf false}, if object arrives at the end of file
            \end{list}

      \end{itemize}

\item Methods returning information about current element of sequence:

      \begin{itemize}
      \item {\em const char* name()} - returns the name of current element (being file or subdirectory)
      \item {\em int fsize()} - returns a size of file or 0 if current element is subdirectory 
      \item {\em bool is\_directory()} - returns {\bf true}, if current element is subdirectory and {\bf false} otherwise
      \item {\em bool is\_hidden()} - returns {\bf true}, if current element is hidden
            (this may be under Windows) and {\bf false} otherwise
      \end{itemize}
      Attention: if the object arrives at the end of file, then values returned by these methods
      are undefined.

\end{enumerate}

\paragraph{Note:}
\begin{list}{}{}
\item Two instances of DirectoryEntry is assigned be equil, if both are:
      \begin{enumerate}
      \item not positioned (the end of files are achieved)
      \item positioned at the same element of the same directory
            supposing that the element is identified by name.
      \end{enumerate}
\end{list}

\subsection{Class DirectoryContainer}

\subsubsection{General information}

DirectoryContainer constitutes a class which provide facilities to work with
file directory %check this turn of speech is correct?
like STL-container.
The value type of DirectoryContainer is class DirectoryEntry,
the iterator type is class DirectoryIterator (see below \ref{DirectoryIterator})
being Forward Iterator for sequence of DirectoryEntry.
In that context DirectoryEntry is considered to be a mean attendant upon fixed record of directory,
positioning of this object by own means must not be fulfill. \\

\noindent Example:

\begin{verbatim}

try{

  DirectoryContainer dir(".");

  DirectoryContainer::iterator it;

  for( it=dir.begin(); it!=dir.end(); ++it ) cout<< it->name() <<endl;

} catch(DirectoryException& ex) {

  cerr << ex.message << endl;
}

\end{verbatim}
This code area has same functionality as that is typed above  

\subsubsection{API}

\paragraph{Associated types:}

\begin{itemize}
\item value\_type - class DirectoryEntry
\item iterator - class DirectoryIterator (\ref{DirectoryIterator}),
\item const\_iterator - the same as iterator
\item reference - a reference to the DirectoryEntry
\item const\_reference - a constant reference to the DirectoryEntry
\item pointer - a pointer to the DirectoryEntry
\item distance\_type - DirectoryIterator difference type (long)
\item size\_type - long
\end{itemize}

\paragraph{Description of methods:}

\begin{enumerate}
\item Container's means:

      \begin{itemize}
      \item {\em DirectoryContainer(DirectoryContainer\&)};
      \item {\em DirectoryContainer\& operator=(DirectoryContainer\&)};
      \item {\em DirectoryIterator begin()};
      \item {\em static const DirectoryIterator\& end()};
      \item {\em size\_type size()};
      \item {\em size\_type max\_size()};
      \item {\em bool empty()};
      \item {\em void swap(DirectoryContainer\&)};
      \end{itemize}

\item Specific means:

      \begin{itemize}
      \item {\em DirectoryContainer()} - creates DirectoryContainer
            associated with current directory;
            {\bf Parameter:} the name of directory;
      \item {\em DirectoryContainer(const char *path)} - creates DirectoryContainer
            associated with given directory;
            {\bf Parameter:} the name of directory;
      \item {\em setPath(const char *path)} - associate DirectoryContainer
            with given directory;
            {\bf Parameter:} the name of directory;
      \item {\em record\_count()} - returns a record count of the directory
            excepting pointer to the directory itself and pointer to the parent direectory;
            it is a size()-2 in most cases
      \item {\em bool no\_records()} - returns {\bf true}, if the directory is not contain any records
            excepting pointer to the directory itself and pointer to the parent direectory,
            and {\bf false} otherwize.

      \end{itemize}

\end{enumerate}


\noindent {\bf Note:}

\begin{list}{}{}
\item Note a size() method returns the record count including special entry
      (which is a pointer to the directory itself and a pointer to the parent direectory),
      and, consequently, an empty() method returns {\bf false} in most cases
      (excepting case of checking empty disk for Windows platform).
      Use no\_files() method to check if the directory is empty in usual sense terms.
\end{list}

\subsection{Class DirectoryIterator} \label{DirectoryIterator}

DirectoryIterator is especial class being a Forward Iterator for sequence of DirectoryEntry.
Thus, it has next methods:
      \begin{itemize}

        \item DirectoryIterator(const DirectoryIterator\& );

        \item DirectoryIterator\& operator=(const DirectoryIterator\& );

        \item DirectoryIterator\& operator++();

        \item DirectoryIterator operator++(int);

        \item DirectoryEntry\& operator*();

        \item const DirectoryEntry\& operator*() const;

        \item DirectoryEntry* operator->();

        \item const DirectoryEntry* operator->() const;

        \item bool operator==(const DirectoryIterator\& );

        \item bool operator!=(const DirectoryIterator\& );

      \end{itemize}
and, then, a defaut constructor DirectoryIterator()
which makes iterator for past-the-end element.

\subsection{Class DirectoryException}

DirectoryException is a class which may be thrown by DirectoryEntry methods
\begin{list}{}{}
\item DirectoryEntry(const char* path);
\item \verb|~|DirectoryEntry();
\item DirectoryEntry(const DirectoryEntry\& x);
\item DirectoryEntry\& operator=(const DirectoryEntry\& x);
\item bool next();
\item bool find(const char* name);
\item void close();
\end{list}
in the case of system error occur, and additionally by {\em DirectoryEntry(cont *char)} constructor
in the case of directory with name being taken is not exist.
As DirectoryIterator and DirectoryContainer are superstructures under DirectoryEntry,
this class DirectoryException may be thrown by next DirectoryIterator's methods:
\begin{list}{}{}
\item DirectoryIterator(const DirectoryIterator\& );
\item operator=(const DirectoryIterator\& );
\item operator++();
\item operator++(int);
\end{list}
and next DirectoryContainer's methods: 
\begin{list}{}{}
\item begin()
\item record\_count()
\item no\_files()
\item size()
\item empty()
\end{list}


\paragraph{API:}

\begin{verbatim}
struct DirectoryException
{
  long errno; // system error number
  string message; // system error message
}

\end{verbatim}

 
  
\section{Full-fledged example}

  This code is full-fledged example of using of the package:

% Данный пример демонстрирует порядок использования пакета {\bf DirectoryContainer}:

\begin{verbatim}
#include <GradSoft/DirectoryContainer.h>

ifdef HAVE_NAMESPACES
#include <iostream>
using namespace std;
using namespace GradSoft;
#else
#include <iostream.h>
#endif

#include <stdio.h>

int main(int argc, char** argv)

  if (argc!=2) {
    cerr << "Usage: " << argv[0] << " path" << endl;
    return 1;
  }

  cout<<endl<<">>>> 1 >>>>"<<endl;
  
  // Usinig DirectoryEntry:
  
  try{
  
    DirectoryEntry smth(argv[1]);  
  
    do {
  
     cout<<smth.name()<<endl;  
  
    } while(smth.next());  
  
  } catch(DirectoryContainer& ex) {
  
    cerr << ex.message << endl;
  }
  
  cout<<endl<<">>>> 2 >>>>"<<endl;
  
  // Using DirectoryContainer:
  
  try{
  
    DirectoryContainer dir(argv[1]);
  
    Directory::iterator it;
  
    for( it=dir.begin(); it!=dir.end(); ++it)
    {
      cout << it->name() ;
      if ( it->is_directory() ) cout << " ( DIR ) ";
      else cout << " ( " << it->fsize() << " ) ";
      cout << endl;
    }    
  
  } catch(DirectoryException& ex) {
  
    cerr << ex.message << endl;
  }
  return 0;
}

\end{verbatim}



\section{ Programming environment conventions }

Using DirectoryContainer on Windows NT, you must:
\begin{enumerate}
  \item to define WIN32 macro before inclusion of DirectoryContainer.h header file;
  \item to use iostream, fstream, etc. standard headers instead iostream.h, fstream.h, etc. ones.
\end{enumerate}

\section{ Changes }

\begin{itemize}
  \item[16-10-2002] fixed incorrect URL for API reference.
  \item[03-01-2002] updated in accordance with 1.4.0 public release.
  \item[01-08-2001] last touch before 1.3.0 public release.
  \item[03-07-2001] name changed to DirectoryContainer
  \item[17-05-2001] bringing into accordance with changes in code.
  \item[11-12-2000] initial revision of english text.
  \item[15-01-2001] name changed to DirectoryIterator.
\end{itemize}

\end{document}
